\toolnameend: A Tool for Visualizing Patterns of Spreadsheet Function Combinations

### 2016-08-13 doc/paper.tex
hrefs

\usepackage{xspace}
\usepackage{subfigure}
\usepackage{url}

@@

	In Excel, the \texttt{IF} functions is one of the logical functions: given a condition
	as the first argument that evaluates to either true or false, it takes on the
	value of either the second or third argument respectively. Or, as the
	documentation writes out, ``\texttt{IF(Something is True, then do something, otherwise
	do something
	else)}"\footnote{\url{https://support.office.com/en-us/article/IF-function-69aed7c9-4e8a-4755-a9bc-aa8bbff73be2?ui=en-US\&rs=en-US\&ad=US}}. Furthermore, the third argument is optional; if left out, the function will return a default ``FALSE" value if necessary. Examples of this function in practice can be seen in Figure~\ref{fig:ifexample}.

@@

	\subsection{Implementation Details} The visualization is the product of two processes: \begin{itemize} \item \textit{Collection}: Given a set of
		Excel sheets, the tool, written in Java, uses Apache
		POI\footnote{https://poi.apache.org/} to identify and iterate over every cell
		containing a valid formula. Afterward, it calls POI's formula parser to decompose
		the formula text into an ordered set of individual tokens, which the tool then
		parses into the tree-like form. When all formulae have been analyzed like this,
		it produces JSON files for each top-level function in the set.

		\item \textit{Presentation}: The JSON files, meanwhile, feed into the
		presentation code, implemented in Javascript with much help from the
		visualization library D3\footnote{https://d3js.org/}. We chose D3 primarily
		for its accessibility: given the JSON, the visualization can be simply
		embedded into a webpage accessible through a browser. \end{itemize} A
	demonstration of \toolname can be found at
	https://github.com/DeveloperLiberationFront/Excel-Function-Visualizer.

@@

	%Lookup example
	For a deeper example, the LOOKUP functions are some of the most popular in
	Excel yet come with a number of important but silent requirements for proper
	use, and, as such, they have warranted some extra attention with regards to
	spreadsheet design. In short, these functions are for ``when you need to look
	in a single row or column and find a value from the same position in a second
	row or
	column"\footnote{\url{https://support.office.com/en-us/article/LOOKUP-function-446d94af-663b-451d-8251-369d5e3864cb}}. Examples of their use can be found in Figure~\ref{fig:lookupexample}, and we will cover each type in more depth to better understand the visualizations for these functions.

@@

	LOOKUP has two forms: vector and array. In either case, the first argument is
	the value which you are trying to match in another column. If the formula is in
	the vector form, that column is the second argument, so in
	Figure~\ref{fig:lookupexample}, the formula in cell D2 (\texttt{=LOOKUP(3,
		A2:A6, B2:B6)}) will search for the value of 3 between cells A2 and A6. The
	value that the formula returns then comes from the corresponding cell in the
	optional third argument's column; since the lookup values matches the value in
	cell A4, it returns the value from cell B4. The array form of the LOOKUP works
	the same way, except that the second argument is a two-dimensional table where
	it matches values in the table's first column and returns values from the last.
	Also, depending on the dimensions of the lookup table, the function may
	substitute rows for columns.

	VLOOKUP works closely with this array form, except the user defines in its
	third argument which column from which to return values. In \texttt{=VLOOKUP(3,
		A2:B6, 2)}, the argument ``2" tells to return the value from the second column,
	which is \textit{B2:B6} here. Furthermore, it also accepts a boolean value for
	an optional argument: when set to FALSE, it returns only exact matches, and
	when TRUE or left out, it approximates the match to the closest value beneath
	it. We demonstrate this in Figure~\ref{fig:lookupexample} in the bottom two
	\texttt{VLOOKUP} functions, one returning ``Friday" and the other an error. \texttt{HLOOKUP}
	works identically but focuses on rows instead of columns.

	All of these variations may seem daunting to a beginner, but \toolname can
	help. Because of its design, users can distinguish between forms of the same
	function, as seen in Figure~\ref{fig:lookups} where one can explore both the
	frequency with which the spreadsheet programmers used these different forms --
	three-argument vector form much more than the array form -- and then retrieve
	examples of every configuration in the dataset. The same goes for the V/HLOOKUP
	functions, as shown partially in Figure~\ref{fig:vhlookups}.

@@

	For these user and case studies, we relied solely on one type of spreadsheet (Excel)
	from one source (Enron), but we have room to develop in either dimension. For sources, 
	Enron's focus on finances and energy represents one context in
	which people use spreadsheets, but other corpora, like EUSES and Fuse, draw from
	other domains and industries. As Jansen found in his comparison of
	spreadsheets Enron and EUSES \cite{jansen2015enron}, different companies rely
	on different functions. For other types of spreadsheet, such as Gnumeric\footnote{http://www.gnumeric.org/} or Google
	Sheets\footnote{https://www.google.com/sheets/about/}, we have no
	guarantee that practices and functions will hold across these varied programs, even within
	similar domains. This, then, warrants further analysis, in which good visualization might serve
	well.

@@

	This material is based upon work supported in whole or in part with funding
	from the Laboratory for Analytic Sciences (LAS). Any opinions, findings,
	conclusions, or recommendations expressed in this material are those of the
	author(s) and do not necessarily reflect the views of the LAS and/or any agency
	or entity of the United States Government. I also thank the anonymous reviewers
	who advised the direction of this paper.

>>>>

\usepackage{xspace}
\usepackage{subfigure}
\usepackage{url}
\usepackage{hyperref}

@@

	In Excel, the \texttt{IF} functions is one of the logical functions: given a condition
	as the first argument that evaluates to either true or false, it takes on the
	value of either the second or third argument respectively. Or, as the
	documentation writes out, ``\texttt{IF(Something is True, then do something, otherwise
	do something
	else)}"\footnote{\href{https://support.office.com/en-us/article/IF-function-69aed7c9-4e8a-4755-a9bc-aa8bbff73be2}{https://support.office.com/en-us/article/IF-function-69aed7c9-4e8a-4755-a9bc-aa8bbff73be2}}. Furthermore, the third argument is optional; if left out, the function will return a default ``FALSE" value if necessary. Examples of this function in practice can be seen in Figure~\ref{fig:ifexample}.

@@

	\subsection{Implementation Details} The visualization is the product of two processes: \begin{itemize} \item \textit{Collection}: Given a set of
		Excel sheets, the tool, written in Java, uses Apache
		POI\footnote{\href{https://poi.apache.org/}{https://poi.apache.org/}} to identify and iterate over every cell
		containing a valid formula. Afterward, it calls POI's formula parser to decompose
		the formula text into an ordered set of individual tokens, which the tool then
		parses into the tree-like form. When all formulae have been analyzed like this,
		it produces JSON files for each top-level function in the set.

		\item \textit{Presentation}: The JSON files, meanwhile, feed into the
		presentation code, implemented in Javascript with much help from the
		visualization library D3\footnote{\href{https://d3js.org/}{https://d3js.org/}}. We chose D3 primarily
		for its accessibility: given the JSON, the visualization can be simply
		embedded into a webpage accessible through a browser. \end{itemize} A
	demonstration of \toolname can be found at
	\href{https://github.com/DeveloperLiberationFront/Excel-Function-Visualizer}{https://github.com/DeveloperLiberationFront/Excel-Function-Visualizer}.

@@

	%Lookup example
	For a deeper example, the LOOKUP functions are some of the most popular in
	Excel yet come with a number of important but silent requirements for proper
	use, and, as such, they have warranted some extra attention with regards to
	spreadsheet design. In short, these functions are for ``when you need to look
	in a single row or column and find a value from the same position in a second
	row or
	column"\footnote{\href{https://support.office.com/en-us/article/LOOKUP-function-446d94af-663b-451d-8251-369d5e3864cb}{https://support.office.com/en-us/article/LOOKUP-function-446d94af-663b-451d-8251-369d5e3864cb}}. Examples of their use can be found in Figure~\ref{fig:lookupexample}, and we will cover each type in more depth to better understand the visualizations for these functions.

@@

	\texttt{LOOKUP} has two forms: vector and array. In either case, the first argument is
	the value which you are trying to match in another column. If the formula is in
	the vector form, that column is the second argument, so in
	Figure~\ref{fig:lookupexample}, the formula in cell D2 (\texttt{=LOOKUP(3,
		A2:A6, B2:B6)}) will search for the value of 3 between cells A2 and A6. The
	value that the formula returns then comes from the corresponding cell in the
	optional third argument's column; since the lookup values matches the value in
	cell A4, it returns the value from cell B4. The array form of the LOOKUP works
	the same way, except that the second argument is a two-dimensional table where
	it matches values in the table's first column and returns values from the last.
	Also, depending on the dimensions of the lookup table, the function may
	substitute rows for columns.

	\texttt{VLOOKUP} works closely with this array form, except the user defines in its
	third argument which column from which to return values. In \texttt{=VLOOKUP(3,
		A2:B6, 2)}, the argument ``2" tells to return the value from the second column,
	which is \textit{B2:B6} here. Furthermore, it also accepts a boolean value for
	an optional argument: when set to FALSE, it returns only exact matches, and
	when TRUE or left out, it approximates the match to the closest value beneath
	it. We demonstrate this in Figure~\ref{fig:lookupexample} in the bottom two
	\texttt{VLOOKUP} functions, one returning ``Friday" and the other an error. \texttt{HLOOKUP}
	works identically but focuses on rows instead of columns.

	All of these variations may seem daunting to a beginner, but \toolname can
	help. Because of its design, users can distinguish between forms of the same
	function, as seen in Figure~\ref{fig:lookups} where one can explore both the
	frequency with which the spreadsheet programmers used these different forms --
	three-argument vector form much more than the array form -- and then retrieve
	examples of every configuration in the dataset. The same goes for the \texttt{V/HLOOKUP}
	functions, as shown partially in Figure~\ref{fig:vhlookups}.

@@

	For these user and case studies, we relied solely on one type of spreadsheet (Excel)
	from one source (Enron), but we have room to develop in either dimension. For sources, 
	Enron's focus on finances and energy represents one context in
	which people use spreadsheets, but other corpora, like EUSES and Fuse, draw from
	other domains and industries. As Jansen found in his comparison of
	spreadsheets Enron and EUSES \cite{jansen2015enron}, different companies rely
	on different functions. For other types of spreadsheet, such as Gnumeric\footnote{\href{http://www.gnumeric.org/}{http://www.gnumeric.org/}} or Google
	Sheets\footnote{\href{https://www.google.com/sheets/about/}{https://www.google.com/sheets/about/}}, we have no
	guarantee that practices and functions will hold across these varied programs, even within
	similar domains. This, then, warrants further analysis, in which good visualization might serve
	well.

@@

	This material is based upon work supported in whole or in part with funding
	from the Laboratory for Analytic Sciences (LAS). Any opinions, findings,
	conclusions, or recommendations expressed in this material are those of the
	author(s) and do not necessarily reflect the views of the LAS and/or any agency
	or entity of the United States Government. We also thank the anonymous reviewers
	who advised the direction of this paper, the study participants for their comments,
	and Shivom Gargava for his help.
